% =============================================================================
% SECTION 2: SYSTEM ARCHITECTURE OVERVIEW
% =============================================================================

\section{System Architecture Overview}
\label{sec:arch}

Figure~\ref{fig:arch} summarizes the proposed architecture.
The design follows three architectural patterns common in modern AI systems~\cite{lewis2020rag}:

\textbf{(1) Retrieval-Augmented Generation (RAG)}---agents retrieve relevant facts from the KG before generating outputs, grounding decisions in structured knowledge rather than parametric memory alone.

\textbf{(2) Tool-Augmented Agents}---LLM agents invoke deterministic tools (optimizer, exporter) for tasks requiring provable correctness~\cite{schick2023toolformer}.

\textbf{(3) Blackboard Architecture}---the KG serves as a shared workspace where specialized agents read inputs, write inferences, and coordinate asynchronously.

The core design principle is that the system is \emph{KG-centered}: extracted plan facts and inferred construction intent are continuously grounded against a structured graph of rules, components, and historical builds.

\begin{figure}[ht]
\centering
\begin{tikzpicture}[
  node distance=4mm and 6mm,
  box/.style={draw, rounded corners, align=center, inner sep=3pt, font=\tiny, minimum width=15mm, minimum height=8mm},
  pde/.style={draw, rounded corners, align=center, inner sep=3pt, font=\tiny, minimum width=15mm, minimum height=8mm, fill=blue!10},
  arrow/.style={-Latex}
]
\node[box] (ingest) {Plan Ingest\\(PDF/CAD)};
\node[box, right=of ingest] (extract) {Geometry +\\Text Extract};
\node[box, right=of extract] (facts) {Plan Facts\\(walls, dims)};
\node[box, below=6mm of facts] (kg) {\textbf{Knowledge}\\textbf{Graph}};
\node[pde, below=6mm of kg] (pde) {Structural\\Solver (PDE)};
\node[box, right=of facts] (agents) {Agent\\Orchestrator};
\node[box, right=of agents] (outputs) {Outputs\\(BOM, Cuts)};

\draw[arrow] (ingest) -- (extract);
\draw[arrow] (extract) -- (facts);
\draw[arrow] (facts) -- (agents);
\draw[arrow] (kg) -- (agents);
\draw[arrow] (kg) -- (pde);
\draw[arrow] (pde) -- (agents);
\draw[arrow] (facts) -- (kg);
\draw[arrow] (agents) -- (outputs);
\end{tikzpicture}
\caption{KG-centered architecture with physics-based structural evaluation: plan facts are extracted, normalized, evaluated using continuum mechanics (PDE solver), and used by agents to generate compliant, optimized outputs.}
\label{fig:arch}
\end{figure}

\subsection{Key Modules}

\textbf{(1) Plan Ingest + Extraction.}
Convert drawings into machine-usable primitives:
vector geometry (lines, polylines), detected room/wall boundaries, and text tokens (dimensions, callouts, schedules)~\cite{pizarro2022floorplan_survey}.
The current implementation supports PDF (via PyMuPDF~\cite{pymupdf2024}), DXF (via ezdxf~\cite{ezdxf2024}), and DWG (via LibreDWG conversion).
For raster images, YOLOv8 object detection~\cite{jocher2023yolov8} and EasyOCR~\cite{easyocr2024} provide robust extraction.

\textbf{(2) Plan Facts Model.}
Normalize extraction into a consistent intermediate representation:
\emph{WallSegments}, \emph{Openings}, \emph{Headers}, \emph{Notes}, \emph{Schedules}, and \emph{Constraints}
with explicit provenance back to the sheet, region, and confidence score.

\textbf{(3) Knowledge Graph.}
Store rules, materials, codes, and historical knowledge (Section~\ref{sec:kg}).
The KG also stores structural hypotheses, beam evaluations, and uncertainty quantification results.

\textbf{(4) Thermo-Hygro-Mechanical Solver (Coupled PDEs).}
The solver evaluates headers, joists, and beams under both \emph{instantaneous} and \emph{long-term} climate exposure.
It couples (i) heat diffusion, (ii) moisture diffusion, and (iii) time-dependent beam mechanics:
\begin{align}
\rho c \, \frac{\partial T}{\partial t} &= \nabla \cdot (k_T \nabla T), \label{eq:heat}\\
\frac{\partial MC}{\partial t} &= \nabla \cdot (D_m \nabla MC), \label{eq:moisture}\\
E_{\mathrm{eff}}(T,MC,t)\, I(t)\, \frac{\partial^4 w(x,t)}{\partial x^4} &= q(x,t). \label{eq:beam_time}
\end{align}
Here $T$ is temperature, $MC$ is moisture content, $w$ is deflection, and $q$ is the distributed load derived from tributary areas and environmental actions.
Climate effects enter through (a) time-varying loads and (b) material evolution:
\begin{align}
q(x,t) &= q_G(x) + q_Q(x,t) + q_M(x,t), \label{eq:q_time}\\
E_{\mathrm{eff}}(T,MC,t) &= \frac{E_0\, f_T(T)\, f_{MC}(MC)}{1+\phi(t,MC)}, \label{eq:Eeff}\\
I(t) &= I_0 \big(1-d(t)\big). \label{eq:It}
\end{align}
$q_G$ is dead load, $q_Q$ includes live and environmental loads (e.g., rain, wind), and $q_M$ captures moisture-related mass (e.g., absorbed water, wet insulation).
$E_{\mathrm{eff}}$ is an effective modulus incorporating temperature/moisture dependence and creep via $\phi(t,MC)$ (viscoelastic + mechano-sorptive),
and $I(t)$ optionally accounts for section loss or degradation through a damage variable $d(t)$ under prolonged high moisture.

The solver outputs time-indexed stress/deflection trajectories and a time-dependent serviceability failure probability:
\begin{equation}
P_{\mathrm{fail}}(t) = \mathbb{P}\!\left(w_{\max}(t) > w_{\mathrm{allow}}\right),
\label{eq:Pfail_t}
\end{equation}
which supports climate-aware ranking of structural hypotheses.

\textbf{(5) Agent Orchestrator.}
A multi-agent workflow that queries the KG, performs reasoning and validation,
invokes the thermo-hygro-mechanical solver for physics-based evaluation over time,
and calls deterministic optimization routines for cutting and procurement.

\textbf{(6) Export + Integration.}
Produce both human-usable artifacts (PDF/HTML summaries, build steps)
and machine-readable exports (CSV/JSON/EDI-style mappings) suitable for lumber yard systems.
Exported structural hypotheses include ranked load path alternatives with tradeoff explanations, including long-term climate resilience.

\textbf{(5) Agent Orchestrator.}
A multi-agent workflow that queries the KG, performs reasoning and validation,
invokes the structural solver for physics-based evaluation,
and calls deterministic optimization routines for cutting and procurement.

\textbf{(6) Export + Integration.}
Produce both human-usable artifacts (PDF/HTML summaries, build steps)
and machine-readable exports (CSV/JSON/EDI-style mappings) suitable for lumber yard systems.
Exported structural hypotheses include ranked load path alternatives with tradeoff explanations.

\subsection{Data Flow}

The system operates in five sequential phases:

\textbf{Phase 1: Extraction.}
Plan documents are parsed into vector geometry, raster images, and text.
YOLOv8 detects symbols; PyMuPDF extracts text and vectors from PDFs; ezdxf handles CAD formats.
Output: raw geometry primitives and text tokens with bounding boxes.

\textbf{Phase 2: Normalization.}
Raw extraction is normalized into \texttt{PlanFact} nodes in the KG, each with source provenance (sheet, region, confidence).
Cross-validation checks dimension consistency and flags ambiguities.

\textbf{Phase 3: Inference.}
Agents query the KG for plan facts and rules, then write inferred assemblies (\texttt{AssemblyIntent}, \texttt{Component}) back to the KG.
Each inference links to triggering facts and applied rules.

\textbf{Phase 3.5: Structural Hypothesis Evaluation.}
The Structural Hypothesis Agent generates multiple plausible vertical load path configurations (typically 3--5 hypotheses), each specifying which walls are load-bearing, where beams or posts are required, and how tributary loads are distributed.
For each hypothesis, headers and beams are evaluated using Euler--Bernoulli beam theory:
\[
EI \frac{d^4 w(x)}{dx^4} = q(x)
\]
where $w(x)$ is deflection, $E$ is modulus of elasticity, $I$ is second moment of area, and $q(x)$ is the distributed load derived from tributary areas.
The solver computes maximum bending stress, shear stress, and deflection, rejecting assemblies that violate mechanical limits.
Uncertainty in tributary widths and material properties is propagated via Monte Carlo sampling, yielding probability-of-failure estimates for each member.
Hypotheses are ranked by cost proxy, structural robustness (1 -- $P_{\text{failure}}$), and design flexibility, with the top 2--3 Pareto-optimal solutions stored in the KG.

\textbf{Phase 4: Optimization.}
The Procurement Agent selects stock items and invokes the cut-stock optimizer (ILP solver) for waste minimization.
Results are written as \texttt{CutPiece} nodes linked to source components.

\textbf{Phase 5: Export.}
Final outputs (BOM, cut lists, instructions) are generated by querying the KG and formatting results.
Every output line item carries provenance links back through phases 4→3→2→1.
