% =============================================================================
% SECTION 5: AGENTIC WORKFLOW
% =============================================================================

\section{Agentic Workflow}
\label{sec:agents}

Agents are responsible for \emph{turning facts into decisions}, while deterministic tools handle optimization and exporting~\cite{wang2024agents_survey, xi2023agents_rise, li2025agentic_survey}.
This separation ensures that reasoning is auditable and optimization is provably correct.

\subsection{Design Principles}

Our agent architecture follows three key principles from recent surveys on tool-augmented LLMs~\cite{zhang2025tool_learning, schick2023toolformer}:

\textbf{Reasoning-Action Separation.}
Following the ReAct pattern~\cite{yao2023react}, each agent interleaves reasoning traces with tool invocations.
Reasoning steps query the KG for facts and rules; actions invoke deterministic tools (optimizer, exporter) or update the KG with validated inferences.
This enables auditability (decisions trace to KG facts), correctness (optimization uses proven algorithms), and debuggability (failed inferences can be replayed).

\textbf{Specialization over Generalization.}
Rather than a single general-purpose agent, we deploy five specialized agents, each with domain-specific prompts, KG query templates, and tool access.
Specialization improves accuracy by constraining the reasoning space.

\textbf{Shared Memory via KG.}
Agents communicate through the knowledge graph rather than direct message passing.
Each agent reads from and writes to the KG, creating a persistent, queryable record of all inferences.

\subsection{Agent Architecture}

Figure~\ref{fig:agents} shows the five agents and their workflow.
Agents (blue) perform reasoning; tools (green) perform deterministic computation.

\begin{figure}[ht]
\centering
\begin{tikzpicture}[
  node distance=5mm and 12mm,
  agent/.style={draw, rounded corners, align=center, inner sep=3pt, font=\scriptsize, fill=blue!10, minimum width=22mm},
  tool/.style={draw, align=center, inner sep=3pt, font=\scriptsize, fill=green!10, minimum width=18mm},
  arrow/.style={-Latex}
]
\node[agent] (qa) {Extraction QA\\Agent};
\node[agent, right=of qa] (infer) {Component\\Inference Agent};
\node[agent, right=of infer] (code) {Code \&\\Compliance Agent};
\node[agent, below=8mm of qa] (procure) {Procurement \&\\Cut Agent};
\node[agent, right=of procure] (instruct) {Instruction\\Generation Agent};
\node[tool, right=of instruct] (optimizer) {Cut-Stock\\Optimizer};

\draw[arrow] (qa) -- (infer);
\draw[arrow] (infer) -- (code);
\draw[arrow] (code.south) -- ++(0,-4mm) -| (procure.north);
\draw[arrow] (procure) -- (instruct);
\draw[arrow] (procure) -- (optimizer);
\end{tikzpicture}
\caption{Agent workflow: specialized agents (blue) handle reasoning; deterministic tools (green) handle optimization. All agents read/write to the shared KG.}
\label{fig:agents}
\end{figure}

\subsection{Inter-Agent Communication}

Agents communicate via the KG using a publish-subscribe pattern:
(1) upstream agent writes inference results as new KG nodes;
(2) downstream agent queries for nodes matching its input criteria;
(3) status tracking via \texttt{pending}/\texttt{validated}/\texttt{flagged} properties;
(4) provenance links to creating agent and reasoning trace.
This decouples agents temporally and logically, enabling parallel execution and full audit trails.

\subsection{Error Handling}

Agents handle uncertainty through tiered strategies:
\textbf{Confidence thresholds}---low-confidence results are flagged for human review rather than propagated.
\textbf{Validation gates}---outputs are checked against schema constraints before KG writes.
\textbf{Graceful degradation}---failed agents mark assemblies as ``requires review'' and continue processing unaffected portions.

\subsection{Extraction QA Agent}

Validates extracted geometry/text, flags low-confidence dimensions:
\begin{itemize}
  \item Dimension consistency checks (wall lengths match scale)
  \item Missing element detection (walls without openings)
  \item OCR confidence thresholds (request higher DPI if $<$0.85)
\end{itemize}

\subsection{Component Inference Agent}

Maps validated plan facts to assemblies using KG rules:
\begin{itemize}
  \item Wall type $\rightarrow$ stud size/spacing, plates, sheathing
  \item Opening size $\rightarrow$ header assembly, jack/king studs
  \item Bearing vs. non-bearing $\rightarrow$ header sizing requirements
\end{itemize}

\subsection{Code \& Compliance Agent}

Checks assemblies against KG code rules~\cite{zhang2023kg_codes, beach2020irc_automation}:
\begin{itemize}
  \item Header sizing for span/load per IRC~\cite{irc2021}
  \item Stud spacing for load-bearing walls
  \item Fastener patterns; produces ``compliant'' or ``requires review'' with citations
\end{itemize}

\subsection{Procurement \& Cut Agent}

Selects stock and invokes the cut optimizer:
\begin{itemize}
  \item Stock lengths (8/10/12/14/16/20 ft, LVL, panels)
  \item Kerf allowance (0.125'') and waste factors
  \item Substitution rules (prefer in-stock; longer stock if less waste)
\end{itemize}

\subsection{Instruction Generation Agent}

Produces build instructions with full traceability:
\begin{itemize}
  \item Organized by wall line, room, or phase
  \item Component references with KG node IDs
  \item Fastener specs and inspection callouts with code citations
\end{itemize}
