% =============================================================================
% SECTION 8: PHASED DEVELOPMENT PLAN
% =============================================================================

\section{Phased Development Plan}

Development follows an iterative, test-driven approach common in AI system engineering~\cite{chase2022langchain}.
Each phase delivers a working system with measurable improvements over the baseline.

\subsection{Phase 1: KG Foundation + Enhanced Takeoff (4--6 weeks)}

\begin{itemize}
  \item Define initial ontology in Neo4j (entities, relationships)
  \item Migrate existing lumber calculations to KG-backed rules
  \item Load baseline component definitions and framing rules
  \item Parse 2--3 plan sets with provenance tracking
  \item Generate BOM with rule citations
\end{itemize}

\textbf{Success Criteria:}
BOM accuracy $>$90\% on validation set; all BOM items link to plan facts and rules; KG query latency $<$100ms.

\subsection{Phase 2: Assembly Reasoning + Cut Optimization (6--10 weeks)}

\begin{itemize}
  \item Expand component library (headers, stud packs, plates, sheathing)
  \item Implement Component Inference Agent with KG queries
  \item Integrate OR-Tools cut-stock optimizer~\cite{ortools2024}
  \item Produce supplier-ready cut list exports
  \item Validate against ground truth projects
\end{itemize}

\textbf{Success Criteria:}
Waste reduction to $<$8\% (baseline: 15--20\%); agent inference accuracy $>$85\%; cut optimization runtime $<$30s per project.

\subsection{Phase 3: Code/Regulation Layer (6--10 weeks)}

\begin{itemize}
  \item Model IRC/IBC code rules with jurisdiction and version~\cite{irc2021}
  \item Implement Code \& Compliance Agent following NLP extraction methods~\cite{zhang2023kg_codes}
  \item Add compliance citations to outputs
  \item Build code rule editor for jurisdiction customization
\end{itemize}

\textbf{Success Criteria:}
Code compliance validation accuracy $>$90\%; all non-compliant assemblies flagged with specific IRC citations; support for 3+ jurisdictions.

\subsection{Phase 4: Feedback Loop + Learning (Ongoing)}

\begin{itemize}
  \item Capture user overrides and corrections
  \item Store outcomes (actual waste, substitutions made)
  \item Continuously improve confidence calibration
  \item Expand KG with new precedents using continuous learning approaches~\cite{zhu2024llm_kg_construction}
\end{itemize}

\textbf{Success Criteria:}
System accuracy improves $>$2\% per 100 projects processed; user override rate decreases over time; KG grows with validated domain knowledge.

\subsection{Risk Mitigation}

\textbf{Data quality:} Start with high-quality architectural plans; flag low-confidence extractions for human review.
\textbf{Code complexity:} Begin with core IRC framing rules; expand incrementally to specialty cases.
\textbf{Integration:} Design JSON/CSV exports to match common supplier formats; iterate based on partner feedback.
